### 2023-08-01 ms.tex

\documentclass[a4paper, 10pt, twocolumn]{article}

\usepackage{graphicx}                % optional für Grafiken
\usepackage[font=footnotesize]{caption}
\usepackage[section]{placeins}
%\usepackage{tabularx}               % optional für Tabellen
%\usepackage{multirow}               % optional für Tabellen
\usepackage{url}                	 % optional für Internet Links
\usepackage{hyperref}

%\usepackage[small,bf]{caption3}     % Bildunterschriften

%\usepackage{parskip}

% verbatim with line line wrapping
% \usepackage{listings}
% \lstset{
%    breaklines=true,
%    basicstyle=\ttfamily}

\usepackage{titlesec}
\usepackage{amsmath}                
\usepackage{amssymb}
\usepackage{abstract}

% temporal packages
\usepackage{lipsum}
\SetLipsumDefault{1}

%\usepackage{enumitem}
%\setlist{nosep} % or \setlist{noitemsep} to leave space around whole list

\usepackage{xcolor}
\newcommand{\highlight}[1]{\colorbox{yellow}{\parbox{\columnwidth}{#1}}}

%\pagestyle{empty}                   % weder Kopf- noch Fußzeile auf 1. Seite
\graphicspath{{figures/}}

\begin{document}

\date{\small\today}                                        % kein Datum auf 1. Seite
\title{
	\vspace{-20mm}\textbf{	\large 
	Phytoplankton Retention Mechanisms in Estuaries: A Case Study of the Elbe Estuary
	}
	}
\author{
	Laurin Steidle\(^1\), Ross Vennel, Jana Hinners \\
	\emph{
		\small
		\(^1\)laurin.steidle@uni-hamburg.de} \\
	\emph{
		\small Universitaet Hamburg, Grosse Elbstrasse 133, Germany}
		}

%--------------------------------------Abstract------------------------------------------------

\twocolumn[
	\begin{@twocolumnfalse}
		\maketitle

		\begin{abstract}
			\begin{center}
				\texttt{All texts in this typewriter font are drafting comments.}
				\smallskip
			\end{center}

			Phytoplankton are essential for estuarine ecosystems.
			However, little is known how estuarine plankton retain themselves in an environment that continuously washes them into unfavorable marine waters.
			We present a new individual-based Lagrangian model of the Elbe estuary examining several retention mechanisms - most importantly vertical migration and tidal stranding.
			We further examine their ability to outgrow losses and provide first estimates on outwashing losses.
			We found that shallow waters are crucial for the long term for the survival of the phytoplankton population.
			\texttt{I feel like it sounds a bit dry. Maybe sparkle in a bit more excitement?}

			\vspace{1cm}
		\end{abstract}
  	\end{@twocolumnfalse}
]

\input{main_text.tex}

\bibliography{references.bib}
\bibliographystyle{unsrt}

\end{document}